### 2019-06-07 1-Preamble/5-Abstract/abstract.tex
Lots of small changes
- added a title/author etc in metadata
- cleaner small proof with an environment
Typo

% ************************** Thesis Abstract *****************************
% Use `abstract' as an option in the document class to print only the titlepage and the abstract.
\begin{abstract}
% \vspace*{1cm}

% \TODOL{Raccourcir juste un peu ?}

% ----------------------------------------------------------------------------

% Chapitre 1
Dans cette thèse de doctorat, nous étudions les réseaux sans fil et les appareils reconfigurables qui peuvent accéder à des réseaux de radio intelligente, dans des bandes non licenciées et sans supervision centrale.
Plus spécifiquement, nous considérons des réseaux de l'Internet des Objets (IoT), avec l'objectif d'augmenter la durée de vie de la batterie des appareils, en les équipant d'algorithmes d'apprentissage machine peu coûteux mais efficaces, qui leur permettent d'améliorer automatiquement l'efficacité de leurs communications sans fil.
%  (Chapitre~\ref{chapter:1}).
% Chapitre 4
Nous proposons différents modèles de réseaux IoT, et nous montrons empiriquement, par des simulations numériques et une validation expérimentale réaliste, le gain que peuvent apporter nos méthodes, qui utilisent l'apprentissage par renforcement.
%  (Chapitre~\ref{chapitre:4}).
% Chapitre 2
Les différents problèmes d'accès au réseau sont modélisés avec des Bandits Multi-Armés (MAB), mais leur analyse est difficile à réaliser,
% , Chapitre~\ref{chapitre:2}
car il est difficile de prouver la convergence de nombreux appareils jouant à un jeu collaboratif sans communication aucune coordination, lorsque les appareils suivent tous un modèle d'activation aléatoire.
% En effet, même le nombre de périphériques actifs à chaque instant est hautement imprévisible, rendant l'environnement non pas stationnaire et évoluant rapidement.
% Chapitres 5 et 6
Le reste de ce manuscrit étudie donc deux modèles restreints, d'abord des bandits multi-joueurs dans des problèmes stationnaires, puis des bandits simple-joueurs non stationnaires.
%  (Chapitre~\ref{chapitre:5})
%  (Chapitre~\ref{chapitre:6})
% Chapitre 3
Nous détaillons également une autre contribution, la bibliothèque Python open-source SMPyBandits pour des simulations numériques de problèmes MAB, qui couvre tous les modèles étudiés et d'autres.
% (Chapitre~\ref{chapter:3})

% \newpage
% ----------------------------------------------------------------------------
\begin{center}
	\textbf{{\Large Abstract}}
\end{center}
% \vspace*{1cm}

% Chapter 1
In this PhD thesis, we study wireless networks and reconfigurable end-devices that can access large-scale Cognitive Radio networks, in unlicensed bands and without central control.
More specifically, we focus on Internet of Things networks (IoT), with the objective of extending the devices' battery life, by equipping them with low-cost but efficient machine learning algorithms, to let them automatically improve the efficiency of their wireless communications.
%  (Chapter~\ref{chapter:1})
% Chapter 4
We propose different models of IoT networks, and we show empirically on both numerical simulations and real-world validation the possible gain of our methods, that use Reinforcement Learning.
%  (Chapter~\ref{chapter:4})
% Chapter 2
The different network access problems are modeled as Multi-Armed Bandits (MAB), but we found that analyzing the realistic models was intractable,
% ,  Chapter~\ref{chapter:2}
because proving the convergence of many end-devices playing a collaborative game without communication no coordination is hard, when end-devices all follow random active pattern.
% Indeed, even the number of active devices at any time is highly unpredictable, making the environment not stationary and quickly evolving.
% Chapter 5 and 6
The rest of this manuscript thus studies two restricted models, first multi-player bandits in stationary problems, then non-stationary single-player bandits.
%  (Chapter~\ref{chapter:5})
%  (Chapter~\ref{chapter:6})
% Chapter 3
We also detail another contribution, the SMPyBandits open-source Python library for numerical simulations of MAB problems, which covers all the studied models and more.
%  (Chapter~\ref{chapter:3})

>>>>

% ************************** Thesis Abstract *****************************
% Use `abstract' as an option in the document class to print only the titlepage and the abstract.
\begin{abstract}
\vspace*{-30pt}
%
% \TODOL{Raccourcir juste un peu ?}
% ----------------------------------------------------------------------------
\begin{small}
% Chapitre 1
Dans cette thèse de doctorat, nous étudions les réseaux sans fil et les appareils reconfigurables qui peuvent accéder à des réseaux de radio intelligente, dans des bandes non licenciées et sans supervision centrale.
Nous considérons des réseaux de l'Internet des Objets (IoT), avec l'objectif d'augmenter la durée de vie de la batterie des appareils, en les équipant d'algorithmes d'apprentissage machine peu coûteux mais efficaces, qui leur permettent d'améliorer automatiquement l'efficacité de leurs communications sans fil.
%  (Chapitre~\ref{chapter:1}).
% Chapitre 4
Nous proposons différents modèles de réseaux IoT, et nous montrons empiriquement, par des simulations numériques et une validation expérimentale réaliste, le gain que peuvent apporter nos méthodes, qui utilisent l'apprentissage par renforcement.
%  (Chapitre~\ref{chapitre:4}).
% Chapitre 2
Les différents problèmes d'accès au réseau sont modélisés avec des Bandits Multi-Armés (MAB), mais leur analyse est difficile à réaliser,
% , Chapitre~\ref{chapitre:2}
car il est difficile de prouver la convergence de nombreux appareils jouant à un jeu collaboratif sans communication aucune coordination, lorsque les appareils suivent tous un modèle d'activation aléatoire.
% En effet, même le nombre de périphériques actifs à chaque instant est hautement imprévisible, rendant l'environnement non pas stationnaire et évoluant rapidement.
% Chapitres 5 et 6
Le reste de ce manuscrit étudie donc deux modèles restreints, d'abord des bandits multi-joueurs dans des problèmes stationnaires, puis des bandits simple-joueurs non stationnaires.
%  (Chapitre~\ref{chapitre:5})
%  (Chapitre~\ref{chapitre:6})
% Chapitre 3
Nous détaillons également une autre contribution, la bibliothèque Python open-source SMPyBandits pour des simulations numériques de problèmes MAB, qui couvre tous les modèles étudiés et d'autres.
% (Chapitre~\ref{chapter:3})
\end{small}

% ----------------------------------------------------------------------------
\hr{}

% \newpage
% ----------------------------------------------------------------------------
\begin{center}
	\textbf{{\Large Abstract}}
\end{center}
% \vspace*{1cm}
% \vspace*{-5pt}
%
\begin{small}
% Chapter 1
In this PhD thesis, we study wireless networks and reconfigurable end-devices that can access large-scale Cognitive Radio networks, in unlicensed bands and without central control.
We focus on Internet of Things networks (IoT), with the objective of extending the devices' battery life, by equipping them with low-cost but efficient machine learning algorithms, to let them automatically improve the efficiency of their wireless communications.
%  (Chapter~\ref{chapter:1})
% Chapter 4
We propose different models of IoT networks, and we show empirically on both numerical simulations and real-world validation the possible gain of our methods, that use Reinforcement Learning.
%  (Chapter~\ref{chapter:4})
% Chapter 2
The different network access problems are modeled as Multi-Armed Bandits (MAB), but we found that analyzing the realistic models was intractable,
% ,  Chapter~\ref{chapter:2}
because proving the convergence of many end-devices playing a collaborative game without communication no coordination is hard, when end-devices all follow random active pattern.
% Indeed, even the number of active devices at any time is highly unpredictable, making the environment not stationary and quickly evolving.
% Chapter 5 and 6
The rest of this manuscript thus studies two restricted models, first multi-player bandits in stationary problems, then non-stationary single-player bandits.
%  (Chapter~\ref{chapter:5})
%  (Chapter~\ref{chapter:6})
% Chapter 3
We also detail another contribution, the SMPyBandits open-source Python library for numerical simulations of MAB problems, which covers all the studied models and more.
%  (Chapter~\ref{chapter:3})
\end{small}

### 2019-06-07 1-Preamble/5-Abstract/resume_fr.tex
Typo

Cette thèse aborde tous les points susmentionnés et propose différentes solutions pour chaque scénario.
Les contributions importantes sont résumées ci-dessous :
\begin{itemize}
\item Item 1
\item Item 2
\item Item 3
\item Item 4
\end{itemize}

@@

Ce manuscrit est organisé comme suit.
Nous rappelons dans le \textbf{Chapitre \ref{chapter:1}} blabla.

Dans le \textbf{Chapitre \ref{chapter:2}}, blabla.

Dans le \textbf{Chapitre \ref{chapter:3}}, blabla.

Dans le \textbf{Chapitre \ref{chapter:4}}, blabla.

Dans le \textbf{Chapitre \ref{chapter:5}}, blabla.

Dans le \textbf{Chapitre \ref{chapter:5}} blabla.

>>>>

Nous pouvons énumérer les points suivants pour résumer les principales contributions de cette thèse.
Nous signalons les conférences internationales auxquelles ont été publiées certains résultats.

\begin{itemize}
    \item
    Nous présentons les concepts et les notations du modèle des bandits multi-bras, dans le chapitre~\ref{chapter:2} selon un point de vue mathématique formel.
    Nous suivons aussi une approche didactique, puisque nous utilisons en Section~\ref{par:2:interactiveDemoDiscoverMAB} une démonstration interactive en ligne conçue pour permettre à n'importe qui de jouer contre un petit problème de bandit depuis son navigateur.

    \item
    Nous donnons une brève revue de la littérature de recherche concernant les algorithmes stochastiques de bandits en Section~\ref{sec:2:famousMABalgorithms}.

    \item
    Nous présentons le problème du choix de l'algorithme qu'un praticien devrait utiliser, ou de la sélection d'algorithme parmi la riche collection de différents algorithms MAB disponibles, dans la Section~\ref{sec:2:chooseYourPreferredBanditAlgorithm}.
    Nous présentons un algorithme appelé \Aggr{} pour l'agrégation d'algorithmes, comme une solution en ligne au problème de sélection d'algorithmes, et nous montrons par quelques simulations numériques qu'il atteint de bonnes performances empiriques
    (WCNC 2018) \cite{Besson2018WCNC}.

    \item
    Nous avons écrit la plus complète bibliothèque de simulation pour les problèmes MAB, appelée SMPyBandits, qui est écrite en Python et publiée en ligne sous une licence open-source \cite{SMPyBandits,SMPyBanditsJMLR}.
    Nous présentons en détail son architecture et ses fonctionnalités dans le Chapitre~\ref{chapter:3}, ainsi que différents exemples de son utilisation.
    Une documentation complète est disponible en ligne, ainsi que des instructions exhaustives pour reproduire les expériences utilisées dans la suite de cette thèse.

    \item
    Nous proposons différents modèles pour les réseaux IdO (IoT), dans le Chapitre~\ref{chapter:4}, où les appareils dotés de capacités de radio intelligente peuvent implémenter de leur côté des algorithmes MAB, pour augmenter automatiquement la durée de vie de leur batterie. Cela permet à davantage de périphériques d'utiliser le même réseau tout en maintenant une haute qualité de service
    (CROWNCOM 2017, ICT demo 2018, WCNC 2019 et MOTIoN 2019)
    \cite{Bonnefoi17,Besson2019WCNC,Bonnefoi2019WCNC}.

    \item
    Nous avons implémenté une preuve de concept du modèle susmentionné \cite{Besson2018ICT}, et nous la présentons en détail en Section~\ref{sec:4:gnuradio}. Nous avons aussi réalisé une courte vidéo présentant notre démonstration, hébergée sur \texttt{\href{https://youtu.be/HospLNQhcMk}{youtu.be/HospLNQhcMk}}.

    \item
    Le code source des deux contributions susmentionnées utilise GNU Octave ou MATLAB et est publié en ligne, ainsi que des instructions claires pour la reproduction de nos travaux.

    \item
    Nous formalisons le modèle du bandit multi-joueur, pour lequel nous avons introduit trois variantes, dans le Chapitre~\ref{chapter:5}.
    Pour le cas avec information de détection (``sensing''), nous proposons deux nouveaux algorithmes, et nous donnons une analyse pour notre algorithme \MCTopM, qui prouve qu'il est asymptotiquement optimal.
    Nous présentons aussi des expériences numériques approfondies pour montrer qu'il est bien plus efficace que les autres algorithmes de la littérature précédente.
    Notre travail \cite{Besson2018ALT} (ALT 2018) a également contribué à une nouvelle impulsion à la recherche sur les bandits multi-joueurs, car certains travaux de recherche récents se sont construits sur nos résultats.

    \item
    Nous donnons également une revue détaillée de la littérature sur les différentes extensions du modèle MAB multi-joueurs, qui, selon nous, n'a jamais été écrit auparavant.
    % (ALT 2018, et 4-8 travaux inspirés de notre article depuis lors). Etat de l'art avec notre algorithme MCTopM + klUCB pour les bandits multi-joueurs "avec détection", état de l'art empirique avec notre approche simple (mais erronée) "égoïste" en cas de "non détection".

    \item
    Nous présentons ensuite le modèle de bandits multi-bras stationnaire par morceaux, dans le chapitre~\ref{chapter:6}, et une vue détaillée de l'état de l'art de la recherche sur ce modèle \cite{Besson2019GLRT,Besson2019Gretsi} (GRETSI 2019).
    Suite à deux travaux récents, nous proposons un nouvel algorithme activement adaptatif pour ce problème stationnaire par morceaux, \GLRklUCB, qui atteint des performances comparables à l'état de l'art.
    % Revue de littérature sur les modèles et algorithmes non stationnaires, état de l'art de la station stationnaire à la pièce avec notre algorithme, test GLR + klUCB

    \item
    La dernière contribution de cette thèse est une revue de la littérature sur les cas d'utilisation possibles de la technique du doublement successif de l'horizon (``doubling trick'') pour les problèmes de bandits,
    ainsi qu'une analyse unifiée et plus générique de deux familles de doublements.
    Ceci a conduit à l'article \cite{Besson2018DoublingTricks}, qui est rapidement présenté en Annexe~\ref{app:2:DoublingTricks}.
\end{itemize}

@@

Ce manuscrit est organisé comme suit.

\begin{figure}[h!]
    \centering
    \resizebox{0.95\textwidth}{!}{
    \begin{tikzpicture}[>=latex',line join=bevel,scale=2.25]
        %
        \node[align=center] (introduction) at (0,3.25) [rectangle,draw] {Chapitre~\ref{chapter:1}\\Introduction};
        \node[align=center] (chapter2) at (0,2.25) [rectangle,draw] {Chapitre~\ref{chapter:2}\\Modèle bandits multi-bras\\stochastique et stationnaire};
        \node[align=center] (chapter3) at (2.5,2.25) [rectangle,draw] {Chapitre~\ref{chapter:3}\\SMPyBandits : bibliothèque\\de simulations pour MAB};
        \node[align=center] (chapter4) at (-2.5,1) [rectangle,draw] {Chapitre~\ref{chapter:4}\\Deux modèles MAB\\pour les réseaux IoT};
        \node[align=center] (chapter5) at (0,1) [rectangle,draw] {Chapitre~\ref{chapter:5}\\Modèle MAB\\Multi-joueurs};
        \node[align=center] (chapter6) at (2.5,1) [rectangle,draw] {Chapitre~\ref{chapter:6}\\Modèle MAB\\Non-stationnaire};
        \node[align=center] (conclusion) at (0,-0.25) [rectangle,draw] {Chapitre~\ref{chapter:conclusion}\\Conclusion générale};
        %
        \draw [color=black,thick,->] (introduction) to (chapter2);
        \draw [color=black,thick,<->] (chapter2) to (chapter3);
        \draw [color=black,thick,->] (chapter2) to (chapter4);
        \draw [color=black,thick,->] (chapter2) to (chapter5);
        \draw [color=black,densely dotted,<->]   (chapter4) to (chapter5);
        \draw [color=black,densely dotted,->] -| (chapter3) to (chapter6);
        \draw [color=black,densely dotted,<->]   (chapter5) to (chapter6);
        \draw [color=black,thick,->] (chapter2) to (chapter6);
        \draw [color=black,thick,->] (chapter4) to (conclusion);
        \draw [color=black,thick,->] (chapter5) to (conclusion);
        \draw [color=black,thick,->] (chapter6) to (conclusion);
        %
    \end{tikzpicture}
    }
    \caption[Organisation de la thèse : une carte de lecture]{Organisation de la thèse. Cette thèse peut se lire en suivant n'importe quel chemin contenant le Chapitre~\ref{chapter:1}, le Chapitre~\ref{chapter:2}, au moins un des trois Chapitres~\ref{chapter:4}, \ref{chapter:5} et \ref{chapter:6}, et la Conclusion.}
    \label{fig:1:organization_fr}
\end{figure}

L'ordre de lecture du manuscrit peut être n'importe quel chemin, descendant entre l'introduction dans le Chapitre~\ref{chapter:1} actuel, et la conclusion dans le dernier Chapitre~\ref{chapter:conclusion}.
Comme le montre le graphique de la Figure~\ref{fig:1:organization_fr} ci-dessus,
la thèse est organisée en deux parties :

\begin{itemize}
    \item
Dans la Partie~\ref{part:Introduction}, nous commençons par le Chapitre~\ref{chapter:2} où nous présentons les modèles MAB, les concepts et les notations utilisés dans tout ce document. Ce premier chapitre est nécessaire à la lecture du reste du manuscrit.
Par contre le Chapitre~\ref{chapter:3} présente notre bibliothèque de simulations SMPyBandits, et même si les Chapitres~\ref{chapter:2}, \ref{chapter:5} et \ref{chapter:6} utilisent la bibliothèque pour leurs simulations numériques, lire ce chapitre n'est pas obligatoire pour comprendre la suite du document.

    \item
La deuxième Partie~\ref{part:MABIOT} contient ensuite trois chapitres, qui sont inclus à la fois dans l'ordre logique et chronologique, mais peuvent être lus presque indépendamment.
Le Chapitre~\ref{chapter:4} commence par présenter différents modèles de réseaux IdO où les algorithmes MAB ont été utilisés avec succès. Nos deux modèles sont intéressants et proches de la réalité, mais ils se sont révélés trop généraux pour proposer une analyse mathématique de la bonne performance empirique des solutions envisagées.
Pour cette raison, nous affaiblissons les modèles pour le reste du document,
et les deux Chapitres~\ref{chapter:5} et \ref{chapter:6} étudient un modèle intermédiaire, situé entre le modèle MAB stationnaire à un joueur du chapitre~\ref{chapter:2} et les modèles de réseaux IdO du chapitre~\ref{chapter:4}.
\end{itemize}

### 2019-06-07 2-Chapters/1-Chapter/chapter1.tex
Typo

    \item
    We also present the piece-wise stationary MAB model, in Chapter~\ref{chapter:6}, and a detailed literature review of the research on non stationary MAB \cite{Besson2019GLRT,Besson2019Gretsi}.
    Following two recent works, we propose a new actively adaptive algorithm for the piece-wise stationary problem, \GLRklUCB, that achieves state-of-the-art performance.
    % - Literature review on non stationary models and algorithms, state-of-the-art for piece-wise stationary with our algorithm, GLR test + klUCB

    \item
    The last contribution of this thesis is a literature review of the possible use cases of the doubling trick technique for MAB problems,
    and a unified and more generic analysis of two families of doubling tricks.
    This lead to the article \cite{Besson2018DoublingTricks}, which is quickly presented in Appendix~\ref{app:2:DoublingTricks}, but not in the main text of the thesis.
\end{itemize}

@@

In Part~\ref{part:Introduction}, we start by the next Chapter~\ref{chapter:2} where we introduce the MAB models, the concepts and the notations used in all this document, and it is mandatory to read it.
Conversely, Chapter~\ref{chapter:3} presents our simulation library SMPyBandits, and even if Chapters~\ref{chapter:2}, \ref{chapter:5} and \ref{chapter:6} use the library for their numerical simulation, reading this chapter is not required to understand the rest of the document.

Then the second Part~\ref{part:MABIOT} contain three chapters, that are included in both the logical and chronological orders, but can be read almost independently.
Chapter~\ref{chapter:4} starts by presenting different models of IoT networks where MAB algorithms have been used with success. Our two models are interesting and close to reality, but they appeared to be too general to propose a mathematical analysis of the good empirical performance of the considered solutions.
For this reason, we weaken the models for the rest of the document,
and both Chapters~\ref{chapter:5} and \ref{chapter:6} studies an intermediate model, lying between the stationary single-player MAB model from Chapter~\ref{chapter:2} and the IoT networks models from Chapter~\ref{chapter:4}.

@@

\paragraph{Seminars.}
    I gave some presentations in the following events:
    SequeL team seminar at Inria Lille in September and December $2017$, and June $2019$;
    SCEE team seminar at CentraleSupélec, Rennes campus, in October $2017$, February $2018$ and June $2019$;
    as well as
    for the GDR ISIS day held in Issy-les-Moulineaux on November $2017$,
    for the brown-bag seminar at ENSAI in Bruz in January $2018$,
    for the weekly seminar at CMAP lab at École Polytechnique in October $2018$,
    for the weekly seminar of the PANAMA project-team at IRISA / Inria Rennes in May $2019$.

\paragraph{Tutorial.}
    I gave a tutorial on the Julia language, at IETR seminar in Vannes in June $2018$,
    with Pierre Haessig (\texttt{pierreh.eu}), see \href{https://HAL.Inria.fr/cel-01830248}{\texttt{HAL.Inria.fr/cel-01830248}}.

\paragraph{Training.}
    I was also in charge of ``GouTP'' training sessions for about $30$ PhD students at CentraleSupélec Rennes.
    We had a lot of various $1$h training sessions between January $2017$ and June $2019$,
    and I gave about $12$ training sessions, on various topics including Python, \texttt{git}, HAL and arXiv, the Julia language and Bib\TeX{}.

@@

\paragraph{Conferences.}
    I attended the following conferences:
	\emph{International Conference on Communication} ICC (Paris), May $2017$,
    \emph{Conference on Learning Theory} COLT (Amsterdam), July $2017$,
    \emph{Conference on Cognitive Radio Oriented Wireless Networks} CROWNCOM (Lisboa), September $2017$,
    \emph{Conference on Algorithmic Learning Theory} ALT (Lanzarote), April $2018$,
    \emph{Wireless Communication and Networking Conference} WCNC (Barcelona), April $2018$,
	\emph{International Conference on Telecommunication} ICT (Saint-Malo), June $2018$,
    \emph{Wireless Communication and Networking Conference} WCNC (Marrakech), April $2019$,
    \emph{Colloque francophone de traitement du signal et des images} GRETSI (Lille), August $2019$.

\paragraph{Seminars.}
    I attended the following seminars:
	\emph{Workshop Learn with Earning} (Rotterdam), May $2018$,
	\emph{Workshop on Optimization and Learning} (Toulouse), September $2018$,
	\emph{European Workshop on Reinforcement Learning} (Lille), October $2018$,

\paragraph{Responsabilities.}
    I was the president of the PhD Students Association of IETR lab (ADDI, \texttt{addi.asso.insa-rennes.fr}) in $2017$.
    I was notably in charge of organizing the PhD Students day at Rennes, in June $2017$, with about $350$ people, and presentation of a research poster\footnote{~See \href{https://HAL.Inria.fr/hal-02013839}{\texttt{HAL.Inria.fr/hal-02013839}}}
    I also helped for the organization and presented another poster\footnote{~See \href{https://HAL.Inria.fr/hal-02013847}{\texttt{HAL.Inria.fr/hal-02013847}}} at the ``IETR : Interagir Évaluer Transmettre Réunir'' seminar, held in Vannes in June $2018$.

\paragraph{Reviews.}
	for \emph{European Workshop on Reinforcement Learning}\footnote{~EWRL 2018, see \href{https://ewrl.wordpress.com/ewrl14-2018}{\texttt{ewrl.wordpress.com/ewrl14-2018}}} (Lille), in October $2018$, I reviewed $5$ papers.
	I helped colleagues for articles submitted at international conferences \emph{AISTATS} $2019$, \emph{NeurIPS} $2017$ et $2018$, and \emph{COLT} $2018$ and $2019$, for about 10 \emph{reviews}.

\paragraph{System Administrator.}
	maintaining our workstations running GNU/Linux and Windows
	at SCEE team,
	and our cognitive radio test-bed using USRP cards and the GNU Radio software,
	from January $2017$ to September $2018$.

@@

\paragraph{Copyright notice.}
%
This document and the additional resources required to compile it (including \LaTeX{} code, Python snippets, images etc)
are \href{https://github.com/Naereen/phd-thesis/}{publicly published},
under the terms of the open-source \href{https://lbesson.mit-license.org/}{\emph{MIT License}},
online at \href{https://github.com/Naereen/phd-thesis/}{\texttt{GitHub.com/Naereen/phd-thesis/}},

>>>>

    \item
    We also present the piece-wise stationary MAB model, in Chapter~\ref{chapter:6}, and a detailed literature review of the research on non stationary MAB \cite{Besson2019GLRT,Besson2019Gretsi} (GRETSI 2019).
    Following two recent works, we propose a new actively adaptive algorithm for the piece-wise stationary problem, \GLRklUCB, that achieves state-of-the-art performance.
    % - Literature review on non stationary models and algorithms, state-of-the-art for piece-wise stationary with our algorithm, GLR test + klUCB

    \item
    The last contribution of this thesis is a literature review of the possible use cases of the sequentially doubling horizon trick technique for MAB problems,
    and a unified and more generic analysis of two families of doubling tricks.
    This lead to the article \cite{Besson2018DoublingTricks}, which is quickly presented in Appendix~\ref{app:2:DoublingTricks}.
\end{itemize}

@@

\begin{itemize}
    \item
In Part~\ref{part:Introduction}, we start by the next Chapter~\ref{chapter:2} where we introduce the MAB models, the concepts and the notations used in all this document, and it is mandatory to read it.
Conversely, Chapter~\ref{chapter:3} presents our simulation library SMPyBandits, and even if Chapters~\ref{chapter:2}, \ref{chapter:5} and \ref{chapter:6} use the library for their numerical simulation, reading this chapter is not required to understand the rest of the document.

    \item
Then the second Part~\ref{part:MABIOT} contain three chapters, that are included in both the logical and chronological orders, but can be read almost independently.
Chapter~\ref{chapter:4} starts by presenting different models of IoT networks where MAB algorithms have been used with success. Our two models are interesting and close to reality, but they appeared to be too general to propose a mathematical analysis of the good empirical performance of the considered solutions.
For this reason, we weaken the models for the rest of the document,
and both Chapters~\ref{chapter:5} and \ref{chapter:6} studies an intermediate model, lying between the stationary single-player MAB model from Chapter~\ref{chapter:2} and the IoT networks models from Chapter~\ref{chapter:4}.
\end{itemize}

@@

\textbf{Seminars.}
    I gave some presentations in the following events:
    SequeL team seminar at Inria Lille in September and December $2017$, and June $2019$;
    SCEE team seminar at CentraleSupélec, Rennes campus, in October $2017$, February $2018$ and June $2019$;
    as well as
    for the GDR ISIS day held in Issy-les-Moulineaux on November $2017$,
    for the brown-bag seminar at ENSAI in Bruz in January $2018$,
    for the weekly seminar at CMAP lab at École Polytechnique in October $2018$,
    for the weekly seminar of the PANAMA project-team at IRISA / Inria Rennes in May $2019$.

\textbf{Tutorial.}
    I gave a tutorial on the Julia language, at IETR seminar in Vannes in June $2018$,
    with Pierre Haessig (\texttt{pierreh.eu}), see \href{https://HAL.Inria.fr/cel-01830248}{\texttt{HAL.Inria.fr/cel-01830248}}.

\textbf{Training.}
    I was also in charge of ``GouTP'' training sessions for about $30$ PhD students at CentraleSupélec Rennes.
    We had a lot of various $1$h training sessions between January $2017$ and June $2019$,
    and I gave about $12$ training sessions, on various topics including Python, \texttt{git}, HAL and arXiv, the Julia language and Bib\TeX{}.

@@

\textbf{Conferences.}
    I attended the following conferences:
	\emph{International Conference on Communication} ICC (Paris), May $2017$,
    \emph{Conference on Learning Theory} COLT (Amsterdam), July $2017$,
    \emph{Conference on Cognitive Radio Oriented Wireless Networks} CROWNCOM (Lisboa), September $2017$,
    \emph{Conference on Algorithmic Learning Theory} ALT (Lanzarote), April $2018$,
    \emph{Wireless Communication and Networking Conference} WCNC (Barcelona), April $2018$,
	\emph{International Conference on Telecommunication} ICT (Saint-Malo), June $2018$,
    \emph{Wireless Communication and Networking Conference} WCNC (Marrakech), April $2019$,
    \emph{Colloque francophone de traitement du signal et des images} GRETSI (Lille), August $2019$.

\textbf{Seminars.}
    I attended the following seminars:
	\emph{Workshop Learn with Earning} (Rotterdam), May $2018$,
	\emph{Workshop on Optimization and Learning} (Toulouse), September $2018$,
	\emph{European Workshop on Reinforcement Learning} (Lille), October $2018$.

\textbf{Responsabilities.}
    I was the president of the PhD Students Association of IETR lab (ADDI, \texttt{addi.asso.insa-rennes.fr}) in $2017$.
    I was notably in charge of organizing the PhD Students day at Rennes, in June $2017$, with about $350$ people, and presentation of a research poster\footnote{~See \href{https://HAL.Inria.fr/hal-02013839}{\texttt{HAL.Inria.fr/hal-02013839}}}
    I also helped for the organization and presented another poster\footnote{~See \href{https://HAL.Inria.fr/hal-02013847}{\texttt{HAL.Inria.fr/hal-02013847}}} at the ``IETR : Interagir Évaluer Transmettre Réunir'' seminar, held in Vannes in June $2018$.

\textbf{Reviews.}
	for \emph{European Workshop on Reinforcement Learning}\footnote{~EWRL 2018, see \href{https://ewrl.wordpress.com/ewrl14-2018}{\texttt{ewrl.wordpress.com/ewrl14-2018}}} (Lille), in October $2018$, I reviewed $5$ papers.
	I helped colleagues for articles submitted at international conferences \emph{AISTATS} $2019$, \emph{NeurIPS} $2017$ et $2018$, and \emph{COLT} $2018$ and $2019$, for about 10 \emph{reviews}.

\textbf{System Administrator.}
	maintaining our workstations running GNU/Linux and Windows
	at SCEE team,
	and our cognitive radio test-bed using USRP cards and the GNU Radio software,
	from January $2017$ to September $2018$.

@@

\textbf{Copyright notice.}
%
This document and the additional resources required to compile it (including \LaTeX{} code, Python snippets, images etc)
are \href{https://github.com/Naereen/phd-thesis/}{publicly published},
under the terms of the open-source \href{https://lbesson.mit-license.org/}{\emph{MIT License}},
online at \href{https://github.com/Naereen/phd-thesis/}{\texttt{GitHub.com/Naereen/phd-thesis/}}.
